### 2024-01-17 presentation/figures/bifurctree3.tex

\begin{tikzpicture}[scale=1.1]
  \draw (0,0) -- (1.1,1.1);
  \draw[dashed] (1.1,1.1) -- (1.4,1.4);
  \draw (1.4,1.4) -- (1.5,1.5);
  \draw (0.5,0.5) -- (1,0);
  \draw (1,1) -- (2,0);
  \draw (1.5,1.5) -- (3,0);

  \filldraw[black] (0.5,0.5) circle (0.75pt) node[anchor=south east]{\color{blue}$\tilde{p}_{1}$};

  \filldraw[black] (1,1) circle (1.5pt) node[anchor=south east]{\color{red}\Large$\tilde{p}_{2}$};
  \filldraw[black] (1.5,1.5) circle (0.75pt) node[anchor=south east]{\color{green}$\tilde{p}_{N-1}$};

  \draw[black] (0,-0.25) node{$p_1$};
  \draw[black] (1,-0.25) node{$p_2$};

  \draw[red,fill=none](.5,-0.25) ellipse (0.75cm  and 0.25cm);

  \draw[black] (2,-0.25) node{$p_3$};
  \draw[red,fill=none](2,-0.25) circle (0.25);
  \draw[black] (3,-0.25) node{$p_{N}$};
\end{tikzpicture}